---
title: "Dirty Data Project - Task 4 - Halloween Candy"
author: John Wood
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(DataExplorer)
```



```{r}

candy_raw_2015 <- read_excel(here("raw_data/boing-boing-candy-2015.xlsx"))
candy_raw_2016 <- read_excel(here("raw_data/boing-boing-candy-2016.xlsx"))
candy_raw_2017 <- read_excel(here("raw_data/boing-boing-candy-2017.xlsx"))


```



```{r}
# Start tidying 2017 data
# Internal ID only appears in 2017 dataset; drop this.
# Make a new ID for each entry that can be consistent with other datasets
# columns at end are not relevant to the task so drop

candy_2017 <- candy_raw_2017 %>% 
  select(-c("Internal ID", "Q10: DRESS":"Click Coordinates (x, y)")) %>% 
  clean_names() %>% 
  rowid_to_column("id") %>% 
  #add identifier to show record is from 2017 dataset
  mutate(id = str_c("2017_", id))


```

```{r}
# Start to make data into tidy format

# Columns starting "Q6 " can be moved into a new column "candy_choice"
# Values in these columns can be moved into a new column "candy_rating"

candy_2017 <- candy_2017 %>% 
  pivot_longer(cols = "q6_100_grand_bar":"q6_york_peppermint_patties",
               names_to = "candy_choice",
               names_prefix = "q6_",
               values_to = "candy_rating"
               )

# rename columns

candy_2017 <- candy_2017 %>% 
  rename(
         trick_or_treat = q1_going_out,
         gender = q2_gender,
         age = q3_age,
         country = q4_country,
         state_province = q5_state_province_county_etc,
         joy_other = q7_joy_other,
         despair_other = q8_despair_other,
         other_comments = q9_other_comments
         )

# add "year" variable to enable grouping by year once datasets are joined
# reorder columns to allow join via bind_rows

candy_2017 <-  candy_2017 %>% 
  mutate(year = 2017) %>% 
  select(sort(names(.)))
```


```{r}
# Same process for 2016 data
# Timestamp is not a unique identifier so drop
# Make a new ID for each entry that can be consistent with other datasets
# End columns not relevant to task so drop

candy_2016 <- candy_raw_2016 %>% 
  select(-c("Timestamp", "Guess the number of mints in my hand.":"[York Peppermint Patties] Ignore")) %>% 
  clean_names() %>% 
  rowid_to_column("id") %>% 
  #add identifier to show record is from 2016 dataset
  mutate(id = str_c("2016_", id))
```


```{r}
# Start to make data into tidy format

# Columns for each candy can be moved into a new column "candy_choice"
# Values in these columns can be moved into a new column "candy_rating"


candy_2016 <- candy_2016 %>% 
  pivot_longer(cols = "x100_grand_bar":"york_peppermint_patties",
               names_to = "candy_choice",
               values_to = "candy_rating"
               )

# rename columns

candy_2016 <- candy_2016 %>% 
  rename(
         trick_or_treat = are_you_going_actually_going_trick_or_treating_yourself,
         gender = your_gender,
         age = how_old_are_you,
         country = which_country_do_you_live_in,
         state_province = which_state_province_county_do_you_live_in,
         joy_other = please_list_any_items_not_included_above_that_give_you_joy,
         despair_other = please_list_any_items_not_included_above_that_give_you_despair,
         other_comments = please_leave_any_witty_snarky_or_thoughtful_remarks_or_comments_regarding_your_choices
         )

# add "year" variable to enable grouping by year once datasets are joined
# reorder columns to allow join via bind_rows

candy_2016 <-  candy_2016 %>% 
  mutate(year = 2016) %>% 
  select(sort(names(.)))
```



```{r}
# Same process for 2015 data
# Timestamp is not a unique identifier so drop
# Make a new ID for each entry that can be consistent with other datasets
# End columns not relevant to task so drop


candy_2015 <- candy_raw_2015 %>% 
  select(-c("Timestamp", "Guess the number of mints in my hand.":"Please estimate the degrees of separation you have from the following folks [Beyoncé Knowles]")) %>% 
  clean_names() %>% 
  rowid_to_column("id") %>% 
  #add identifier to show record is from 2015 dataset
  mutate(id = str_c("2015_", id))
```

```{r}
# Start to make data into tidy format

# Columns for each candy can be moved into a new column "candy_choice"
# Values in these columns can be moved into a new column "candy_rating"



candy_2015 <- candy_2015 %>% 
  pivot_longer(cols = "butterfinger":"york_peppermint_patties",
               names_to = "candy_choice",
               values_to = "candy_rating"
               )


# rename columns

candy_2015 <- candy_2015 %>% 
  rename(
         trick_or_treat = are_you_going_actually_going_trick_or_treating_yourself,
         age = how_old_are_you,
         joy_other = please_list_any_items_not_included_above_that_give_you_joy,
         despair_other = please_list_any_items_not_included_above_that_give_you_despair,
         other_comments = please_leave_any_remarks_or_comments_regarding_your_choices
         )

# add "year" variable to enable grouping by year once datasets are joined
# also add missing columns with value NA to ensure join is ok
# reorder columns to allow join via bind_rows

candy_2015 <-  candy_2015 %>% 
  mutate(year = 2015,
         gender = NA,
         country = NA,
         state_province = NA) %>% 
  select(sort(names(.)))

```


```{r}
# now join all three datasets together by binding columns

candy_all <- bind_rows(candy_2015, candy_2016, candy_2017)

# reorder columns for ease of visual inspection

candy_all <- candy_all %>% 
  relocate(id, age, gender, trick_or_treat, candy_choice, candy_rating,
          joy_other, despair_other, other_comments, country, state_province, year)

# can now start trying to clean the data!
```

```{r}
# Recode candy_choice entries
# No easy way of doing this using regex so has to be coded manually :-(
# Not changing joke entries/long names, only potential duplicates/misspellings


candy_all <- candy_all %>% 
  mutate(candy_choice = case_when(
                  candy_choice %in% c("anonymous_brown_globs_that_come_in_black_and_orange_wrappers",
                                      "anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes")
                                                                         ~ "mary_janes",
                  candy_choice %in% c("bonkers", "bonkers_the_candy")    ~ "bonkers",
                  candy_choice == "boxo_raisins"                         ~ "box_o_raisins",
                  candy_choice == "dark_chocolate_hershey"               ~ "hersheys_dark_chocolate",
                  candy_choice %in% c("licorice_not_black", "licorice_yes_black")
                                                                         ~ "licorice",
                  candy_choice == "sweetums_a_friend_to_diabetes"        ~ "sweetums",  #When you think of Pawnee, you think of Sweetums
                  candy_choice == "x100_grand_bar"                       ~ "100_grand_bar",
                  TRUE                                                   ~ candy_choice
                  )
         )
            
```

```{r}
# Remove joke entries from the candy_choice observations.
# Responses to "any_full_sized_candy" are not informative for assessing individual candy choices
# Therefor I have decided to remove this option from the analysis.

candy_all <- candy_all %>% 
  filter(!candy_choice %in% c("any_full_sized_candy_bar",
          "vials_of_pure_high_fructose_corn_syrup_for_main_lining_into_your_vein",
          "candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants",
          "cash_or_other_forms_of_legal_tender",
          "dental_paraphenalia",
          "generic_brand_acetaminophen",
          "glow_sticks",
          "broken_glow_stick",
          "creepy_religious_comics_chick_tracts",
          "hugs_actual_physical_hugs",
          "lapel_pins",
          "pencils",
          "peterson_brand_sidewalk_chalk",
          "vicodin",
          "white_bread",
          "whole_wheat_anything",
          "bonkers_the_board_game",
          "person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes",
          "real_housewives_of_orange_county_season_9_blue_ray"
          ))


```

ANALYSIS
# 1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}

candy_all %>% 
  distinct(candy_rating)

# A candy rating occurs when the column candy_rating is populated
# Four possible entries: "JOY", "DESPAIR", "MEH" and NA
# So want count of entries that are not NA

candy_all %>% 
  summarise(total_number_of_ratings = sum(!is.na(candy_rating)))
```


ANALYSIS
# Q2 What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

# Answering this first requires age data to be cleaned

```{r}
# Use these to identify what needs cleaning, materiality level of entries still to be cleaned
candy_all %>% 
  distinct(id, .keep_all = TRUE) %>% 
  group_by(age) %>% 
  summarise(age_count = n()) %>% 
  arrange(age)

candy_all %>% 
  distinct(age) %>% 
  arrange(age)

# Clean age values

candy_all <- candy_all %>% 
  #remove any ".0" or other number combos - cut down number of decimals
  mutate(age = str_remove_all(age, "\\.[0-9]*")) %>% 
  # remove character values
  mutate(age = str_remove_all(age, "[a-zA-Z]*")) %>% 
  # remove other punctuation 
  mutate(age = str_remove_all(age, "[ ?!'\\-,>^+():]*")) %>% 
  # now change variable to be a number. Some records will be coerced to NA
  # however won't make a material difference
  mutate(age = as.numeric(age)) %>% 
  # set bounds on age range: 8 - 100
  # if outside these bounds then set to the limit
  # (tried to set to NA but was getting error)
  mutate(age = case_when(
                  age < 8     ~ 8,
                  age <= 100  ~ age,
                  age > 100   ~ 100
         )
         )

```


#q2 What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
candy_all %>% 
  distinct(id, .keep_all = TRUE) %>% 
  group_by(trick_or_treat) %>% 
  summarise(average_age = mean(age, na.rm = TRUE))


```


#q3 For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
candy_all %>% 
  mutate(joy = (candy_rating == "JOY"),
         despair = (candy_rating == "DESPAIR"),
         meh = (candy_rating == "MEH")
         ) %>% 
  group_by(candy_choice) %>% 
  summarise(total_joy = sum(joy, na.rm = TRUE),
            total_despair = sum(despair, na.rm = TRUE),
            total_meh = sum(meh, na.rm = TRUE)
            ) %>% 
  slice_max(total_joy, n = 5) %>% 
  slice_max(total_despair, n = 5) %>% 
  slice_max(total_meh, n = 5)
 
 #NEED TO REVISIT THIS 
  
```



How many people rated Starburst as despair

```{r}

candy_all %>% 
  filter(candy_choice == "starburst" & candy_rating == "DESPAIR") %>% 
  summarise(total_starburst_despair = n())
```
```{r}
# investigate free text entries in despair_other to find any mention of starburst 

candy_all %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(id, despair_other) %>% 
  filter(str_detect(despair_other,"starburst"))

# only one record returned, so not worth looking at further
```


For the next three questions, count despair as -1, joy as +1 and meh as 0.
```{r}
 #create a new_rating variable using this system:
candy_all <- candy_all %>% 
 
  mutate(new_rating = case_when(
                        candy_rating == "JOY"       ~ 1,
                        candy_rating == "DESPAIR"   ~ -1,
                        candy_rating == "MEH"       ~ 0,
                        TRUE                        ~ 0
                        )
  )
```


What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_all %>% 
  filter(!is.na(gender) & candy_choice != "any_full_sized_candy_bar") %>% 
  group_by(gender, candy_choice) %>% 
  summarise(total_new_rating = sum(new_rating)) %>% 
  slice_max(total_new_rating)
```


What was the most popular candy bar in each year?

```{r}
candy_all %>% 
  group_by(year, candy_choice) %>% 
  summarise(total_new_rating = sum(new_rating)) %>% 
  slice_max(total_new_rating)
```

What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?





```{r}
# Country responses to be cleaned
# Task only asks for US/Canada/UK/Other
# Majority of responses are some variant of US
# Clean these and Canada/UK 
# Assign all other non-NA responses to "Other" - this is accurate enough for the task.

# Send all country results to lower case and remove punctuation 
candy_all <- candy_all %>% 
  mutate(country = str_to_lower(country),
         country = str_remove_all(country, "[:punct:]")) %>% 
  arrange(country)



distinct(candy_all, country)


# Visual inspection of the data shows that all entries beginning with "u' are for USA
# with the exception of "uae", "uk", "united kingdom" and "united kindom"
# First change these to country = "UK". 
candy_all <- candy_all %>% 
  mutate(country = case_when(
              country %in% c("uk", "united kindom", "united kingdom") ~ "UK",
              TRUE ~ country)
         )

# Now convert all entries beginning "u" to "USA"
candy_all <- candy_all %>% 
  mutate(country = case_when(
              str_sub(country, 1, 1) == "u" ~ "USA",
              TRUE ~ country
              )
         )



distinct(candy_all, country)

# Remaining variations of USA:
usa_variations <- c("ahemamerca",
                    "alaska",
                    "america",
                    "california",
                    "i pretend to be from canada but i am really from the united states",
                    "merica",
                    "murica",
                    "murrika",
                    "n america",
                    "new jersey",
                    "new york",
                    "north carolina",
                    "pittsburgh",
                    "subcanadian north america merica",
                    "the best one usa",
                    "the united states",
                    "the united states of america",
                    "the yoo ess of aaayyyyyy"
                    )

candy_all <- candy_all %>% 
  mutate(country = case_when(
                    country %in% usa_variations ~ "USA",
                    TRUE ~ country
                    )
         )
  
# Remaining variants of UK:

uk_variations <- c("endland", "england", "scotland")

candy_all <- candy_all %>% 
  mutate(country = case_when(
                    country %in% uk_variations ~ "UK",
                    TRUE ~ country
                    )
         )

# Variants of Canada:
canada_variations <- c("can", "canada", "canada`", "canae", "soviet canuckistan")

candy_all <- candy_all %>% 
  mutate(country = case_when(
                    country %in% canada_variations ~ "Canada",
                    TRUE ~ country
                    )
         )

# Change all other countries to "Other"

candy_all <- candy_all %>% 
  mutate(country = case_when(
                  country %in% c("Canada", "UK", "USA") ~ country,
                  TRUE ~ "Other"
                  )
         )


```


