---
title: "Dirty Data Project - Task 4 - Halloween Candy"
author: John Wood
output: html_notebook
---

This task involves cleaning and analysis of 3 years data from the BoingBoing Candy survey.
The data is particularly messy, particularly with free-text respon.
The cleaning process is outlined below but the data cleaning script has more detail.

#### Data Formatting

- The task requires joining the datasets for 2015, 2016 and 2017. To enable this, each
of the datasets was cut down to remove unneeded columns, column names were cleaned and
renamed as required, and a unique id added to each record.
- The raw data is in wide-format and was transformed into long-format. Individual columns
for each candy choice and rating were moved into a single column.

#### Data Cleaning
_Candy choice_

- The offered candy choices were slightly different in each year so recoding was done
to capture potential duplicates.
- There were a number of 'joke' entries that were not candy choices. It was decided to 
remove these so that the analysis only considered 'real' candy.
- Additionally, one catch-all choice was "Any Full Sized Candy". It was decided to 
also remove this, as it distorted results on individual candy choices.

_Age_

- Age responses were free-text, so contained text, symbols etc which were removed.

_Country_

- Country was a free-text option. The task asked specific questions about responses from 
USA, UK, Canada and Other Countries so the cleaning focussed on the first three countries. 
Most responses were variants of "USA". 


```{r include = FALSE}
library(tidyverse)
library(here)
```



```{r include = FALSE}

candy_all <- read_csv(here("clean_data/candy_all.csv"))


```


#### ANALYSIS OF DATA

 Q1. What is the total number of candy ratings given across the three years.

```{r}
# A candy rating occurs when the column candy_rating is populated
# Four possible entries: "JOY", "DESPAIR", "MEH" and NA
# So want count of entries that are not NA

candy_all %>% 
  summarise(total_number_of_ratings = sum(!is.na(candy_rating)))
```



 Q2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
candy_all %>% 
  distinct(id, .keep_all = TRUE) %>% 
  group_by(trick_or_treat) %>% 
  summarise(average_age = mean(age, na.rm = TRUE))
```


```{r}
# function to output the top 3 ratings for the sentiment chosen.
most_ratings <- function(sentiment){
  
  candy_all %>% 
    mutate(rating = (candy_rating == sentiment),
           ) %>% 
    group_by(candy_choice) %>% 
   summarise(total = sum(rating, na.rm = TRUE),
              ) %>% 
  slice_max(total, n = 3)
}

```

```{r message = FALSE}
most_ratings("JOY")
most_ratings("DESPAIR")
most_ratings("MEH")
```


Q4. How many people rated Starburst as despair

_This looks at the ratings in the candy_ratings column.
Additional ratings for Starburst may be found in the free text responses. An
initial analysis of these responses found only one record for Starburst, so no
further work was done on this._
```{r}

candy_all %>% 
  filter(candy_choice == "starburst" & candy_rating == "DESPAIR") %>% 
  summarise(total_starburst_despair = n())
```
```{r include = FALSE}
# investigate free text entries in despair_other to find any mention of starburst 

candy_all %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(id, despair_other) %>% 
  filter(str_detect(despair_other,"starburst"))

# only one record returned, so not worth looking at further
```



```{r}
# The next three questions use a rating system of count despair as -1, joy as +1 and meh as 0
# Create a new_rating variable using this system:
candy_all <- candy_all %>% 
 
  mutate(new_rating = case_when(
                        candy_rating == "JOY"       ~ 1,
                        candy_rating == "DESPAIR"   ~ -1,
                        candy_rating == "MEH"       ~ 0,
                        TRUE                        ~ 0
                        )
  )
```


Q5. What was the most popular candy bar by this rating system for each gender in the dataset?
```{r message = FALSE}
candy_all %>% 
  filter(!is.na(gender)) %>% 
  group_by(gender, candy_choice) %>% 
  summarise(total_new_rating = sum(new_rating)) %>% 
  slice_max(total_new_rating)
```

Q6. What was the most popular candy bar in each year?
```{r message = FALSE}
candy_all %>% 
  group_by(year, candy_choice) %>% 
  summarise(total_new_rating = sum(new_rating)) %>% 
  slice_max(total_new_rating)
```
Q7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
```{r message = FALSE}

candy_all %>% 
  group_by(country, candy_choice) %>% 
  summarise(total_new_rating = sum(new_rating)) %>% 
  slice_max(total_new_rating)

```






