---
title: "Dirty Data Project - Task 1 - Decathlon analysis"
author: John Wood
output: html_notebook
---

This task involves cleaning a raw dataset of decathlon event results and performing analysis on it.

The raw dataset consists of one row per athlete, with a series of columns with their result in each event.

## Cleaning script
The script reads in the raw dataset and performs the following actions:

* the row names are converted to a variable "athlete"
* athlete names are converted to consistent case
* clean and rename variable names for clarity
* the data is transformed into a tidy format by assigning the current event columns
  into a single "event" variable with a corresponding "event_result" variable. This is done 
  using the pivot_longer function from the tidyr package.

```{r include = FALSE}
# Load required libraries and read clean data

library(tidyverse)
library(here)

decathlon_clean <- read_csv(here("clean_data/decathlon_clean.csv"))
```


## Question 1:  Who had the longest long jump seen in the data?
```{r }
decathlon_clean %>%
  filter(event == "long_jump") %>% 
  slice_max(event_result)
```


## Question 2: What was the average 100m time in each competition?
```{r }
decathlon_clean %>%
  filter(event == "x100m") %>% 
  group_by(competition) %>% 
  summarise(avg_100m_time = mean(event_result))
```


## Question 3: Who had the highest total points across both competitions?
```{r }
decathlon_clean %>% 
  filter(event == "decathlon_total") %>% 
  group_by(athlete) %>% 
  summarise(total_points = sum(event_result)) %>%
  slice_max(total_points)
```


## Question 4: What were the shot-put scores for the top three competitors in each competition?
_I have interpreted this as the top three competitors overall, i.e. highest total points_
```{r }
decathlon_clean %>%
  filter(event == 'shot_put' & decathlon_rank <= 3)
```


## Question 5: What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r }

decathlon_clean %>% 
  group_by(athlete) %>% 
  # add 400m time category
  mutate(time_for_400m = case_when(
                                  event == "x400m" & event_result <= 50 ~ "Under 50secs",
                                  event == "x400m" & event_result > 50  ~ "Over 50secs"
                                  )
         ) %>% 
  # populate the athlete's time_for_400m for each athlete record
  fill(time_for_400m, .direction = "downup") %>% 
  ungroup() %>% 
  filter(event == "decathlon_total") %>% 
  group_by(time_for_400m) %>% 
  summarise(average_points = mean(event_result))

```




