---
title: "Dirty Data Project - Task 1 - Decathlon analysis"
author: John Wood
output: html_notebook
---

```{r include = FALSE}
# Load required libraries and read clean data

library(tidyverse)
library(here)

decathlon_clean <- read_csv(here("clean_data/decathlon_clean.csv"))
```


## Question 1:  Who had the longest long jump seen in the data?
```{r echo = FALSE}

decathlon_clean %>%
  select(athlete, long_jump, competition) %>% 
  slice_max(long_jump)

```


## Question 2: What was the average 100m time in each competition?
```{r echo = FALSE}

decathlon_clean %>%
  group_by(competition) %>% 
  summarise(avg_100m_time = mean(x100m))

```


## Question 3: Who had the highest total points across both competitions?
```{r echo = FALSE}

decathlon_clean %>% 
  group_by(athlete) %>% 
  summarise(total_points = sum(points)) %>%
  slice_max(total_points)
  
```


## Question 4: What were the shot-put scores for the top three competitors in each competition?
```{r echo = FALSE}

decathlon_clean %>%
  select(competition, athlete, shot_put) %>% 
  group_by(competition) %>% 
  slice_max(shot_put, n = 3)
  
```


## Question 5: What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r echo = FALSE}

decathlon_clean %>%
  mutate(time_for_400m = if_else(x400m <= 50, "Under 50sec", "Over 50sec")) %>% 
  group_by(time_for_400m) %>% 
  summarise(average_points = mean(points))

```

